Loading and preprocessing the data
========================================================

In this section we load the data into R and make transformation on the date column to be read as date and not as factor.

```{r, echo=TRUE}
data<-read.csv("activity.csv",na.strings="NA")  ##1##

data$date<-as.Date(data$date)   ##2##
```

What is mean total number of steps taken per day?
========================================================

1. Plot a histogram of the total number of steps taken in each day.

```{r Histogram, echo=TRUE}
good<-complete.cases(data$steps,data$date) ## Ignore NAs
plotdata<-data[good,]                      ## Ignore NAs

##aggregate based on total steps per date##
totalperday <- aggregate(plotdata$steps,list(Date=plotdata$date),data=plotdata,mean) 

##plot histogram##
hist(totalperday$x,main="Histogram of the total number of steps taken each day",xlab="Number of Steps",col="red")

mean<-mean(totalperday$x)  

median<-median(totalperday$x)
```

2. Then mean of the number of steps taken per day is `r mean` and the median of the steps
taken per day is `r median`.

What is the average daily activity pattern?
================================================

1. Plot number of steps averaged across all days
```{r, echo=TRUE}
good<-complete.cases(data$steps,data$date) ## Ignore NAs
plotdata<-data[good,]                      ## Ignore NAs

##aggregate based on total steps per interval##
total<- aggregate(plotdata$steps,list(Interval=plotdata$interval),data=plotdata,mean) 

##plot graph##
plot(total$Interval,total$x,type="l",ylab="Average Number of Steps",xlab="Time Interval")

##Find maximum time interval
whichmax<-which.max(total$x)
max<-total$Interval[whichmax]
```
2. The interval [`r max`,`r max+5`] contains the maximum number of steps, averaged across all days.

Inputing missing values
===============================

```{r, echo=TRUE}
na<-sum(is.na(data$steps))
```
1. The number of missing values in the dataset is `r na`.

2. If a value for a 5-minute interval is missing we set the value to be the mean of that 5-minute interval averaged across all days.

3. Plot histogram of the total number of steps per day of the new dataset with the filled missing values.


```{r, echo=TRUE}
good<-complete.cases(data$steps,data$date) ## Ignore NAs
plotdata<-data[good,]                      ## Ignore NAs

##aggregate based on total steps per interval##
total<- aggregate(plotdata$steps,list(Interval=plotdata$interval),data=plotdata,mean) 

## create new dataset##
newdata<-data
newdata$steps=NULL

##Fill the NA values##
new<-is.na(data$steps)

for(i in 1:length(data$steps)){
        if(new[i] == "TRUE"){

                if(i %% 288 == 0){newdata$steps[i]=total$x[288]}
                 
                     else {newdata$steps[i]=total$x[i %% 288]}    
        }
        else 
            newdata$steps[i]=data$steps[i]    
}


##aggregate based on total steps per interval##
totalnew<- aggregate(newdata$steps,list(Interval=newdata$interval),data=newdata,mean) 

##Plot the histogram of the new data##
hist(totalnew$x,main="Histogram of the total number of steps taken each day",xlab="Number of Steps",col="red")


newmean<-mean(totalnew$x)  

newmedian<-median(totalnew$x)
```

The new mean of the number of steps taken per day is `r newmean` and the new median of the steps taken per day is `r newmedian`.

 
Are there differences in activity patterns between weekdays and weekends?
=========================================================================

Create a new factor variable in the dataset which indicates whether a day is weekday or weekend.

```{r, echo=TRUE}
## create a new factor in newdata set##
for(i in 1:length(newdata$date)){
        
        if(weekdays(newdata$date[i]) %in% c("Saturday","Sunday")){
                
                newdata$index[i]="weekend"
                
        }
                else{newdata$index[i]="weekday"}
        
}
```
Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, echo=TRUE}
##Find mean of days based on weekday or weekend##
totaldays<- aggregate(newdata$steps,list(days=newdata$index,interval=newdata$interval),data=newdata,mean) 

##Plot steps averaged across weekends or weekdays
library(lattice)
xyplot(x ~ interval | days, data = totaldays,type="l", layout = c(1, 2),xlab="Interval",ylab="Number of Steps")

```
